Regression Cube Analysis of Cohort Study Data

### 2015-06-20 draft.tex
citations are now treated as gramatically invisible

% -- My Own Packages and Commands
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\newcommand{\rem}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\add}[1]{\textcolor{blue}{#1}}
\newcommand{\com}[1]{\textcolor{red}{\underline{#1}}}
\newcommand{\design}[1]{\textcolor{orange}{#1}}
\newcommand{\magenta}[1]{\textcolor{magenta}{#1}}
% white, black, red, green, blue, cyan, magenta, yellow.

@@

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
	%%
  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	%%
  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand.
	%%
  \item An openly available web-based system, which incorporates a open data format to be applicable for data of other application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
\noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%%
\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
%%
We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either regarding a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed w.r.t. different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space.
%%
It also introduces a new information bias, as assumptions are modeled through the discretization.
%%

Medical image data is also acquired and analyzed in modern cohort studies.
%%
We incorporate image-derived features, but do not focus on analyzing image data.
%%
More discussions of cohort study data types and characteristics can be found in \cite{Preim2014, Toennies2015}.
%%
\\\\
Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%%
The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely for older subjects.
%%
It also influences the general body condition and thereby almost all features acquired through cohort studies.
%%
Another important confounder is \emph{gender}.
%%
Other confounders have to be selected by epidemiologists specific to the investigated condition.

@@

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per cohort subject.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
In 1977, Tukey already stated that data is too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
\\\\
The work of Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan}, Angelelli et al.~\cite{Angelelli} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Akin to our approach, Piringer et al.~\cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
%%
We analyze a large amount of simpler models in terms of \emph{different} features.
%%
They also focus on metric features, while we process categorical data as well.
%%

%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} use the same term to describe an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to \cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
Generalized Pairs plots (\texttt{GPLOM'S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
They are therefore suitable for \emph{small} epidemiological data sets.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data is visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical Analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their research agenda acts as guideline for our design.
%%
Namely, comprising perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

@@

%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
%%
They extract classification rules that serve as basis for our proof-of-concept tests.
%%
In \cite{Niemann2015}, Niemann et al. improve the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate findings presented in \cite{Niemann2014} further.
%%
\paragraph{Prior Work.}
In our prior work, we analyzed the healthy aging process of the lumbar spine.
%%
We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
%%
%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%

We defined an Interactive Visual Analysis workflow for image-centric cohort study data \cite{Klemm2014VIS}. 
%%
We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
%%
Information visualizations of cohort study features were augmented with extracted image data, yielding linked views with both image and non-image information.
%%
The mosaic plot presented in \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
%%
Its great popularity among our domain experts was the inspiration for this work.
%%

@@

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes} (based on~\cite{Keim}).
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
\noindent Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

@@

%%
We conducted the analysis with three participants.
%%
\emph{KH}, a clinician (10 years of experience) with focus on epidemiological research, is the domain expert for the breast fat data set analysis.
%%
She is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.
%%
The hepatic steatosis data set is analyzed by \emph{UN}, a data scientists responsible for prior analysis of the data.
%%
The third participant is \emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
%%
\subsection{The Hepatic Steatosis Data Set}
We employ the data set used in \cite{Niemann2014} to identify predictive features w.r.t. the reversible hepatic steatosis disorder.
%%
The dichotomous target feature is derived from the liver fat concentration measured using MRI scans.
%%
Liver fat concentrations of no more than 10\% are mapped to the `negative' class; values greater than 10\% are mapped to the `positive' class to indicate absence or presence of the disease.
%%
The data set contains labels for 578 participants for which the MRI data was available to that point.
%%
The MRI scans for each subject have been introduced in SHIP-2, so the target feature is only present the last study wave.

@@

%%
In \cite{Niemann2014}, the authors show different class distributions of liver fat concentrations of women and men.
%%
For women, an association between age and liver fat was identified.
%%
An appropriate cut-off value of 52 years was set, yielding the most homogeneous class distribution within the resulting subsets.
%%
Based on these observations, we perform our analysis on three populations: \emph{males}, \emph{females (all ages)} and \emph{females older than 52 years}.
%%
\subsection{The Breast Fat Data Set}
The breast fat data set was compiled to find associations between the parenchyma tissue proportion in the female breast compared to other features in the data.
%%
The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
%%
Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.

>>>>

% -- My Own Packages and Commands
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\newcommand{\rem}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\add}[1]{\textcolor{blue}{#1}}
\newcommand{\com}[1]{\textcolor{red}{#1}}
\newcommand{\design}[1]{\textcolor{orange}{#1}}
\newcommand{\magenta}[1]{\textcolor{magenta}{#1}}
% white, black, red, green, blue, cyan, magenta, yellow.

@@

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
	%%
  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	%%
  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand.
	%%
  \item An openly available web-based system, which incorporates a open data format to be applicable for data of other application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
\noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\com{TODO: Reviewer suggested deletion of this sentence, but starting with subsection right after the section is ugly.}
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

We described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step \cite{Klemm2014VIS}.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%%
\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
%%
We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either regarding a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed w.r.t. different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space.
%%
It also introduces a new information bias, as assumptions are modeled through the discretization.
%%

Medical image data are also acquired and analyzed in modern cohort studies.
%%
We incorporate image-derived features, but do not focus on analyzing image data.
%%
More discussions of cohort study data types and characteristics can be found in the works of Preim et al.~\cite{Preim2014} and Toennies et al.~\cite{Toennies2015}.
%%
\\\\
Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%%
The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely for older subjects.
%%
It also influences the general body condition and thereby almost all features acquired through cohort studies.
%%
Another important confounder is \emph{gender}.
%%
Other confounders have to be selected by epidemiologists specific to the investigated condition.

@@

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per cohort subject.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
In 1977, Tukey already stated that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
\\\\
The work of Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan}, Angelelli et al.~\cite{Angelelli} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Akin to our approach, Piringer et al.~\cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
%%
We analyze a large amount of simpler models in terms of \emph{different} features.
%%
They also focus on metric features, while we process categorical data as well.
%%

%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} use the same term to describe an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
Generalized Pairs plots (\texttt{GPLOM'S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
They are therefore suitable for \emph{small} epidemiological data sets.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical Analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their research agenda acts as guideline for our design.
%%
Namely, comprising perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

@@

%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
%%
They extract classification rules that serve as basis for our proof-of-concept tests.
%%
Niemann et al.~\cite{Niemann2015} improved the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
%%
\paragraph{Prior Work.}
In our prior work, we analyzed the healthy aging process of the lumbar spine.
%%
We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
%%
%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%

We defined an Interactive Visual Analysis workflow for image-centric cohort study data \cite{Klemm2014VIS}. 
%%
We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
%%
Information visualizations of cohort study features were augmented with extracted image data, yielding linked views with both image and non-image information.
%%
The mosaic plot \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
%%
Its great popularity among our domain experts was the inspiration for this work.
%%

@@

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes}.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
\noindent Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

@@

%%
We conducted the analysis with three participants.
%%
\emph{KH}, a clinician (10 years of experience) with focus on epidemiological research, is the domain expert for the breast fat data set analysis.
%%
She is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.
%%
The hepatic steatosis data set is analyzed by \emph{UN}, a data scientists responsible for prior analysis of the data.
%%
The third participant is \emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
%%
\subsection{The Hepatic Steatosis Data Set}
We employ the data set used by Niemann et al.~\cite{Niemann2014} to identify predictive features w.r.t. the reversible hepatic steatosis disorder.
%%
The dichotomous target feature is derived from the liver fat concentration measured using MRI scans.
%%
Liver fat concentrations of no more than 10\% are mapped to the `negative' class; values greater than 10\% are mapped to the `positive' class to indicate absence or presence of the disease.
%%
The data set contains labels for 578 participants for which the MRI data was available to that point.
%%
The MRI scans for each subject have been introduced in SHIP-2, so the target feature is only present the last study wave.

@@

%%
Niemann et al.~\cite{Niemann2014} show different class distributions of liver fat concentrations of women and men.
%%
For women, an association between age and liver fat was identified.
%%
An appropriate cut-off value of 52 years was set, yielding the most homogeneous class distribution within the resulting subsets.
%%
Based on these observations, we perform our analysis on three populations: \emph{males}, \emph{females (all ages)} and \emph{females older than 52 years}.
%%
\subsection{The Breast Fat Data Set}
The breast fat data set was compiled to find associations between the parenchyma tissue proportion in the female breast compared to other features in the data.
%%
The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
%%
Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.
